---
title: "Coursera Capstone"
author: "Crisitan Neuhaus"
date: "15/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, include=FALSE}
if(!require(AppliedPredictiveModeling))install.packages("AppliedPredictiveModeling")
if(!require(caret))install.packages("caret")
if(!require(pgmm))install.packages("pgmm")
if(!require(rpart))install.packages("rpart")
if(!require(tidyverse))install.packages("tidyverse")
if(!require(randomForest))install.packages("randomForest")
if(!require(rpart.plot))install.packages("rpart.plot")
if(!require(corrplot))install.packages("corrplot")
if(!require(rattle))install.packages("rattle")
if(!require(tm))install.packages("tm")
if(!require(downloader))install.packages("downloader")
```
Libraries used:  
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(AppliedPredictiveModeling)
library(caret)
library(pgmm)
library(rpart)
library(tidyverse)
library(rpart.plot)
library(randomForest)
library(corrplot)
library(rattle)
library(tm)
library(downloader)
```
# Week #1 {.tabset .tabset-fade} 

## Overview  

Read data from [Capstone data Set](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip)  

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Execute this junk first time to download file and unzip it
if(FALSE){
Url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
download.file(Url,destfile="Coursera-SwiftKey.zip",method="curl")
unzip ("Coursera-SwiftKey.zip", exdir = "./")
}
```

Sampling. To reiterate, to build models you don't need to load in and use all of the data. Often relatively few randomly selected rows or chunks need to be included to get an accurate approximation to results that would be obtained using all the data. Remember your inference class and how a representative sample can be used to infer facts about a population. You might want to create a separate sub-sample dataset by reading in a random subset of the original data and writing it out to a separate file. That way, you can store the sample and not have to recreate it every time. You can use the rbinom function to "flip a biased coin" to determine whether you sample a line of text or not.  

## Q-1  

The *en_US.blogs.txt* file is how many megabytes?  

( ) 150  
( ) 250  
(x) 200  
( ) 100

```{r echo=TRUE, message=FALSE, warning=FALSE}
file.size("./final/en_US/en_US.blogs.txt")/1024/1024
```
## Q-2  

The *en_US.twitter.txt* has how many lines of text?  

( ) Around 5 hundred thousand  
(x) Over 2 million  
( ) Around 1 million  
( ) Around 2 hundred thousand  

```{r echo=TRUE, message=FALSE, warning=FALSE}
# en_US_Data <- read.delim("./final/en_US/en_US.twitter.txt") # too big
con <- file("./final/en_US/en_US.twitter.txt", "r") 
# readLines(con, 1) ## Read the first line of text 
# readLines(con, 1) ## Read the next line of text
# readLines(con, 5) ## Read in the next 5 lines of text

length(readLines(con))

close(con) ## It's important to close the connection when you are done. See the connections help page for more information.
```

## Q-3  

What is the length of the longest line seen in any of the three en_US data sets?  

(x) Over 40 thousand in the blogs data set  
( ) Over 11 thousand in the news data set  
( ) Over 11 thousand in the blogs data set  
( ) Over 40 thousand in the news data set  


```{r echo=TRUE, message=FALSE, warning=FALSE}
files_data <- list("./final/en_US/en_US.twitter.txt",
                   "./final/en_US/en_US.blogs.txt",
                   "./final/en_US/en_US.news.txt")
for(i in 1:length(files_data)){
  con <- file(paste(files_data[i],sep = '"'), "r")
  lenn <- nchar(readLines(con))
  print(paste(files_data[i],"longest line as: ", max(lenn)))
  close(con)
}
```

## Q-4  

In the en_US twitter data set, if you divide the number of lines where the word "love" (all lowercase) occurs by the number of lines the word "hate" (all lowercase) occurs, about what do you get?  

( ) 2  
( ) 0.25  
( ) 0.5  
(x) 4  

```{r echo=TRUE, message=FALSE, warning=FALSE}
con <- file("./final/en_US/en_US.twitter.txt", "r")
love <- grepl(".love.", readLines(con), ignore.case = FALSE)
close(con)
con <- file("./final/en_US/en_US.twitter.txt", "r")
hate <- grepl(".hate.", readLines(con), ignore.case = FALSE)
close(con)
sum(love)/sum(hate)
```

## Q-5  

The one tweet in the en_US twitter data set that matches the word "biostats" says what?  

( ) They just enrolled in a biostat program  
(x) They haven't studied for their biostats exam  
( ) They need biostats help on their project  
( ) It's a tweet about Jeff Leek from one of his students in class  

```{r echo=TRUE, message=FALSE, warning=FALSE}
con <- file("./final/en_US/en_US.twitter.txt", "r")
temp <- readLines(con)
temp[grepl(".biostats.", temp, ignore.case = FALSE) ==TRUE]
close(con)
```

## Q-6

How many tweets have the exact characters "A computer once beat me at chess, but it was no match for me at kickboxing". (I.e. the line matches those characters exactly.)  

( ) 0  
(x) 3  
( ) 2  
( ) 1  

```{r echo=TRUE, message=FALSE, warning=FALSE}
con <- file("./final/en_US/en_US.twitter.txt", "r")
temp <- readLines(con)
temp[grepl("A computer once beat me at chess, but it was no match for me at kickboxing", temp, ignore.case = FALSE) ==TRUE]

close(con)
```

# Week #2 {.tabset .tabset-fade} 

## Task 2  

**Task 2 - Exploratory Data Analysis**  
The first step in building a predictive model for text is understanding the distribution and relationship between the words, tokens, and phrases in the text. The goal of this task is to understand the basic relationships you observe in the data and prepare to build your first linguistic models.  

**Tasks to accomplish**  

Exploratory analysis - perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora.  
Understand frequencies of words and word pairs - build figures and tables to understand variation in the frequencies of words and word pairs in the data.  

**Questions to consider**  

 - Some words are more frequent than others - what are the distributions of word frequencies?  
 - What are the frequencies of 2-grams and 3-grams in the dataset?  
 - How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%?  
 - How do you evaluate how many of the words come from foreign languages?  
 - Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words in the dictionary to cover the same number of phrases?  
 
## Task 3
 
**Task 3 - Modeling**
The goal here is to build your first simple model for the relationship between words. This is the first step in building a predictive text mining application. You will explore simple models and discover more complicated modeling techniques.  

**Tasks to accomplish**  

Build basic n-gram model - using the exploratory analysis you performed, build a basic n-gram model for predicting the next word based on the previous 1, 2, or 3 words.  
Build a model to handle unseen n-grams - in some cases people will want to type a combination of words that does not appear in the corpora. Build a model to handle cases where a particular n-gram isn't observed.  

**Questions to consider** 

 - How can you efficiently store an n-gram model (think Markov Chains)?  
 - How can you use the knowledge about word frequencies to make your model smaller and more efficient?  
 - How many parameters do you need (i.e. how big is n in your n-gram model)?  
 - Can you think of simple ways to "smooth" the probabilities (think about giving all n-grams a non-zero probability even if they aren't observed in the data)?  
 - How do you evaluate whether your model is any good?  
 - How can you use backoff models to estimate the probability of unobserved n-grams?  
 
**Hints, tips, and tricks**  

As you develop your prediction model, two key aspects that you will have to keep in mind are the size and runtime of the algorithm. These are defined as:  

**Size:** the amount of memory (physical RAM) required to run the model in R  
**Runtime:** The amount of time the algorithm takes to make a prediction given the acceptable input
Your goal for this prediction model is to minimize both the size and runtime of the model in order to provide a reasonable experience to the user.  

Keep in mind that currently available predictive text models can run on mobile phones, which typically have limited memory and processing power compared to desktop computers. Therefore, you should consider very carefully (1) how much memory is being used by the objects in your workspace; and (2) how much time it is taking to run your model. Ultimately, your model will need to run in a Shiny app that runs on the shinyapps.io server.  

**Tips, tricks, and hints**  

Here are a few tools that may be of use to you as you work on their algorithm:  

 - **object.size():** this function reports the number of bytes that an R object occupies in memory  
 - **Rprof():** this function runs the profiler in R that can be used to determine where bottlenecks in your function may exist. The profr package (available on CRAN) provides some additional tools for visualizing and summarizing profiling data.  
 - **gc():** this function runs the garbage collector to retrieve unused RAM for R. In the process it tells you how much memory is currently being used by R.  
There will likely be a tradeoff that you have to make in between size and runtime. For example, an algorithm that requires a lot of memory, may run faster, while a slower algorithm may require less memory. You will have to find the right balance between the two in order to provide a good experience to the user.  

## Peer  

**Instructions**  
The goal of this project is just to display that you've gotten used to working with the data and that you are on track to create your prediction algorithm. Please submit a report on R Pubs (http://rpubs.com/) that explains your exploratory analysis and your goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. You should make use of tables and plots to illustrate important summaries of the data set. The motivation for this project is to: 1. Demonstrate that you've downloaded the data and have successfully loaded it in.2. Create a basic report of summary statistics about the data sets.3. Report any interesting findings that you amassed so far.4. Get feedback on your plans for creating a prediction algorithm and Shiny app.  

**Review criteria**  
 - Does the link lead to an HTML page describing the exploratory analysis of the training data set?  
 - Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?  
 - Has the data scientist made basic plots, such as histograms to illustrate features of the data?  
 - Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?  
